TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-06-07 paper/draft.tex
add cannon ref

While elemental abundance measurements from high resolution spectra of F, G, and K stars have achieved extremely high precisions (down to errors in the range of $0.01-0.03$\,dex; \citealt{Nissen2018}) enabled by the improvement of atmospheric models (citations),
the determination of precise metallicities for M dwarfs has remained a long standing challenge, due to the formation of complex molecular features at temperatures, particularly TiO and VO which dominate the optical, H$_2$O and CO which dominate the infrared, and hydride molecules (FeH, CaH, CrH, MgH, etc.) which form at late spectral types \citep{Allard1997}.

@@

In this work we build a data-driven model for M dwarf spectra with \thecannon\ (\citealt{Ness:2015}, \citealt{Ho:2017a}, \citealt{Casey:2016}).
\thecannon\ is a pixel-by-pixel regression for spectral data given a training
set of stars with reliable (assumed) accurate labels.
The model can both predict a spectrum given labels, and also predict labels given a spectrum.
Thus it is a generative model for stellar spectra that can be used for parameter inference.

>>>>

While elemental abundance measurements from high resolution spectra of F, G, and K stars have achieved extremely high precisions (down to errors in the range of $0.01-0.03$\,dex; \citealt{Nissen2018}) enabled by the improvement of atmospheric models (citations),
the determination of precise metallicities for M dwarfs has remained a long standing challenge, due to the formation of complex diatomic and triatomic molecular features at M dwarf temperature regimes, particularly TiO and VO which dominate the optical, H$_2$O and CO which dominate the infrared, and hydride molecules (FeH, CaH, CrH, MgH, etc.) which form at late spectral types \citep{Allard1997}.

@@

In this work we build a data-driven model for M dwarf spectra with \thecannon\ (\citealt{Ness:2015}, \citealt{Ho:2017a}, \citealt{Casey:2016}, \citealt{Behmard2019}).
\thecannon\ is a pixel-by-pixel regression for spectral data given a training
set of stars with reliable (assumed) accurate labels.
The model can both predict a spectrum given labels, and also predict labels given a spectrum.
Thus it is a generative model for stellar spectra that can be used for parameter inference.
